---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Rachel Weiner"
date: "January 5, 2026"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

Based on Mitch Hedberg joke that the hotel chain, "La Quinta" translates to "next to Denny's" as these two locations are commonly clustered together across American towns and cities. Let's see where this joke pans out according to the data. 

### Exercise 1

Let's start by collecting the dimensions of the denny's data set. 

```{r count_dennys}
nrow(dennys)
ncol(dennys)
```

There are 1643 rows in the data set which each represent a different Denny's restaurant location. There are also six columns in this data set which each represent a different variable regarding each Denny's location. The six variables include address, city, state, zip code, longitude and latitude. 

### Exercise 2

Next, let's see about the La Quinta data set. 

```{r count _laquinta}
nrow(laquinta)
ncol(laquinta)
```

There are 909 rows in this data set each representing a different La Quinta hotel location. There are also six columns each representing the same six variables as listed in the Denny's data set. (Address, city, state, zip code, longitude, and latitude)

### Exercise 3

There are many La Quinta hotels outside of the United States including Canada, Mexico, China, New Zealand, Georgia, and Turkiye. This information is clearly listed on their website.

On the otherhand, Denny's does not clearly list any restauraunts as being outside of the United States and makes clear that they are "America's Diner, In your Neighborhood" suggesting that they only have United States locations. They might have locations that are outside of the United States but this is not made clear via their website. 

### Exercise 4

As the answers above were based on each of the organization's websites, the data will provide a more substantive answer to the question regarding whether these organizations have locations outside of the United States. 

The first brainstormed idea that I came up with works with the state variable. I might be able to filter the data to reveal rows that did not have a state mentioned. This might help me to visualize which locations, of both companies, were not located within the United States. 

### Exercise 5

Using the filter() function, I will find whether there are any Denny's locations outside of the United States. 

```{r filter_variable}
dennys %>%
  filter(!(state %in% states$abbreviation))
```

Based off of this, there are no Denny's locations outside of the United States from this data set. 

### Exercise 6

Now I will add a seventh variable to the Dennys data set to reflect the country in which each restaurant is located, which is the United States for all locations. 

```{r mutate_variable}
dennys <- dennys %>%
  mutate(country = "United States")
```

### Exercise 7


### Exercise 8

```{r}
laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
    state == "AG" ~ "Antigua and Barbuda",
    state == "CH" ~ "Switzerland",
    state == "FM" ~ "Micronesia", 
    state == "PU" ~ "Guinea-Bissau",
    state == "QR" ~ "Qatar",
    state == "SL" ~ "Sierra Leone",
    state == "VE" ~ "Venezuela"))
```


### Exercise 9

### Exercise 10

### Exercise 11

### Exercise 12

